### 2019-04-26 index/00-abstract.Rmd
thesis -> initial release

Bacillus Calmette–Guérin (BCG) remains the only licensed vaccine against Tuberculosis. In 2005, England changed from universal vaccination of school-age children to targeted vaccination of high-risk neonates. Little work has been done to assess the impact of this policy change. This thesis evaluates the impact of this change.  \par

Whilst the characteristics of Tuberculosis in England have been reported elsewhere, little attention has been given to the role of BCG. Consequently, I explored and combined, the available data sources. Reporting on data quality issues, trends in incidence rates and differences in outcomes stratified by BCG status. \par

Prior to the change in policy, several studies were carried out to assess the impact. The strength of this evidence has not been reassessed. I recreated one such study and found that there was a greater impact than previously thought.  \par

Determining the benefits of being BCG vaccinated is necessary to properly assess the impact of the policy change. I evaluated the evidence that vaccination may improve outcomes for Tuberculosis cases in England and found that there was some evidence of an association between vaccination and reduced mortality. \par

Surveillance data can help assess whether changes in vaccination policy have influenced incidence rates. I used surveillance data to determine whether those at school-age, or neonates, were directly affected by the policy change. I found the policy change was associated with increased notifications in the UK born but this was outweighed by a reduction in notifications in the non-UK born. \par

Statistical modelling is restricted by the available data. Incorporating non-linear and indirect effects can also be complex. Therefore, I developed a dynamic model of Tuberculosis, fitting to available data, and forecast the long term and indirect effects of the policy change. I found that school-age vaccination *placeholder* Tuberculosis incidence rates compared with neonatal or no vaccination. \par