### 2019-05-03 rmds/00-acknowledgements.Rmd
reorganised thesis to work with online gitbook
Former-commit-id: 3fd8c59438bc22ef6b09d75159886a79ea855f4e

 Thank you to Ellen Brooks-Pollock and Hannah Christensen, for supervising this thesis and for providing guidance and support. I would also like to Matthew Hickman for providing additional supervisory support. \par

  This thesis could not have been completed with out the support of the Tuberculosis section at Public Health England. Of particular help were: Maeve K Lalor, Dominik Zenner, and Colin Campbell. Mary Ramsay, also at Public Health England, provided invaluable insights. \par

  Nicky Welton provided support with the statistical methodology used in Chapter 7 - thank you. \par

  This work was funded by the National Institute for Health Research Health Protection Research Unit (NIHR HPRU) in Evaluation of Interventions at University of Bristol in partnership with Public Health England. Thank you for this opportunity. \par

  Thank you to my friends and family for providing support over 4 years. \par

  Finally, thank you to my partner, Venexia Walker, without whom this would definitely have gone unfinished.