### 2019-05-03 rmds/00-declaration.Rmd

I declare that the work in this thesis was carried out in accordance with the requirements of the University of Bristol’s Regulations and Code of Practice for Research Degree Programs and that it has not been submitted for any other academic award. Except where indicated by specific reference in the text, the work is my own work. Work done in collaboration with, or with the assistance of other, is indicated as such. Any views expressed in this thesis are those of the author. \par

Signed: \par

Date: *placeholder*