### 2019-12-16 rmds/00-contribution.Rmd
added Sutherland changes

Abbott S. *getTBinR: an R package for accessing and summarising the World Health Organisation Tuberculosis data*,
Journal of Open Source Software, 2019, 4(34), 1260. doi: https://doi.org/10.21105/joss.01260 \par

SA conceived, designed, and undertook the work. SA wrote the paper and approved the work for publication. \par

Abbott S., Christensen H., Brooks-Pollock E. *Reassessing the evidence for universal school-age Bacillus Calmette Guerin (BCG) vaccination in England and Wales*. doi: https://doi.org/10.1101/567511 \par

SA, HC, and EBP conceived and designed the work. SA undertook the analysis with advice from all other authors. All authors contributed to the interpretation of the data. SA wrote the first draft of the paper and all authors contributed to subsequent drafts. All authors approve the work for publication and agree to be accountable for the work. \par

Abbott S., Christensen H., Lalor M. K., Zenner D., Campbell C., Ramsay M. E., & Brooks-Pollock, E. (2019). *Exploring the effects of BCG vaccination in patients diagnosed with tuberculosis: Observational study using the Enhanced Tuberculosis Surveillance system*. Vaccine, 1–6. doi: http://doi.org/10.1016/j.vaccine.2019.06.056 (preprint: https://doi.org/10.1101/366476) \par

SA, HC, and EBP conceived and designed the work. SA undertook the analysis with advice from all other authors. All authors contributed to the interpretation of the data. SA wrote the first draft of the paper and all authors contributed to subsequent drafts. All authors approve the work for publication and agree to be accountable for the work. \par

Abbott S., Christensen H., Welton N.J., Brooks-Pollock E. (2019) *Estimating the effect of the 2005 change in BCG policy in England: a retrospective cohort study, 2000 to 2015*. Eurosurveillance, 49(24). doi: https://doi.org/10.2807/1560-7917.ES.2019.24.49.1900220 \par

SA, HC, and EBP conceived and designed the work. NJW provided guidance on the statistical methods used. SA undertook the analysis with advice from all other authors. All authors contributed to the interpretation of the data. SA wrote the first draft of the paper and all authors contributed to subsequent drafts. All authors approve the work for publication and agree to be accountable for the work. \par

\vspace{10mm}

Signed :  \par

\vspace{10mm}

Signed: \par

\vspace{10mm}

Date: